---
title: "Reproducible Research: Peer Assessment 1"
date: "Feb 5th 2016"
output: 
  knitrBootstrap::bootstrap_document:
    title: "ReproAssess1"
    theme: journal
    highlight: Tomorrow
---

```{r echo=F, warning=FALSE,message=F}
library("dplyr")  
library("ggplot2")
library("tidyr")
```

## Loading and preprocessing the data

```{r}
df<-read.csv("activity.csv")
df=mutate(df,date=as.Date(date, format = '%Y-%m-%d'))
knitr::kable(head(df),align="l")
```

## What is total number of steps taken per day?
```{r}
df.day_sum=df %>% group_by(date) %>% 
  summarize(total_step=sum(steps,na.rm=T),mean_step=mean(steps,na.rm=T),median_step=median(steps,na.rm=T),num_na=sum(is.na(steps))) #%>%
  #filter(num_na>0)
knitr::kable(head(df.day_sum))
ggplot(df.day_sum,aes(total_step))+
  geom_histogram(bins = 10)+
  theme_bw()
```

### mean and median of the total number steps per day
```{r}
knitr::kable(summarise(df.day_sum,mean=mean(total_step),median=median(total_step)),align="l")
```

## What is the average daily activity pattern?
```{r}
df.time_sum=df %>% group_by(interval) %>% 
  summarize(mean_step=mean(steps,na.rm=T),median_step=median(steps,na.rm=T))
knitr::kable(head(df.time_sum))
ggplot(df.time_sum,aes(interval,y=mean_step))+
  geom_point()+
  geom_line()+
  theme_bw()
```

###time interval that has maxium step
```{r}
knitr::kable(filter(df.time_sum,mean_step==max(df.time_sum$mean_step)))
```

## Imputing missing values
###total number of missing value
```{r}
#df.sum_na=data.frame(num_na=sum(is.na(df$steps)),total_num=dim(df)[1])
knitr::kable(summarise(df,num_missing=sum(is.na(steps)),num_total=length(steps)),align="l")
```

###impute missing value using the median of that interval across all days
```{r}
df.imput=df
for (i in unique(df.imput$interval) ){
  value=filter(df.time_sum,interval==i)$median_step[1]
  df.imput=mutate(df.imput,steps=ifelse(interval==i & is.na(steps),value,steps))
  
#  idx=which(df.imput$interval==i & is.na(df.imput$steps))
#  print(i,value,idx)
#  df.imput[idx,]$steps=value
}
df.imput.day_sum=df.imput %>% group_by(date) %>% 
  summarize(total_step=sum(steps,na.rm=T),mean_step=mean(steps),median_step=median(steps),num_na=sum(is.na(steps))) #%>%
  #filter(num_na>0)
knitr::kable(head(df.imput.day_sum))
ggplot(df.imput.day_sum,aes(total_step))+
  geom_histogram(bins=10)+
  theme_bw()
```

### mean and median of the total number steps per day
```{r}
before_impute=summarise(df.day_sum,mean=mean(total_step),median=median(total_step))
after_impute=summarise(df.imput.day_sum,mean=mean(total_step),median=median(total_step))
df.tmp=rbind(before_impute,after_impute)
rownames(df.tmp)=c("before","after")
knitr::kable(df.tmp,align="l")
```

###plot difference in mean value for each day
draw lines between changes before and after NA imputation
```{r}
df.comp.day_sum=merge(df.day_sum,df.imput.day_sum,by="date") %>%
  select(date,before_impute=total_step.x,after_impute=total_step.y)
  
  
df.comp.day_sum.rmna=gather(df.comp.day_sum,key=impute,value=steps,-date) 
#df.comp.day_sum.rmna=mutate(df.comp.day_sum.rmna,mod_steps=ifelse(is.na(steps),-1,steps))

ggplot(df.comp.day_sum.rmna,aes(date,steps,group=date))+
  geom_point(aes(color=impute,size=2,alpha=0.5)) +
  geom_line()
```

###table show dates have changed mean steps
```{r}
knitr::kable(filter(df.comp.day_sum,before_impute!=after_impute | is.na(before_impute)))
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
df.weekday=mutate(df,weekday=ifelse(weekdays(df$date) %in% c("Sunday","Saturday"),"WeekEnd","WeekDays"))
df.weekday_sum=df.weekday %>% group_by(weekday,interval) %>% 
  summarize(mean_step=mean(steps,na.rm=T),median_step=median(steps,na.rm=T))
ggplot(df.weekday_sum,aes(interval,mean_step)) + 
  facet_wrap(~weekday,nrow=2) +
  geom_line() +
  theme_bw()
```

##for double check the result
something doesn't seem right that all 8 dates that have NA value imputed to the same total
```{r}
df.wide=spread(df,date,steps)
knitr::kable(df.wide[1:5,1:10])
```
but after spread the table, 
8 days has totally missing data for 288 interval 
it is clear that for each interval, 
all imputed NA has the same value, which is the median of that interval across all date, 
thus, the total number of steps per day increase the same for these 8 days
the mean number increase 8x1141/62=147.23 days
because they are calculate to be 0 instead of NA before impute
the median

